---
title: "Week 4 Part 1 - CIs and T-tests"
author: "Will Geiken"
date: "10/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## 1. Confidence intervals

```{r}
otter_length <- c(38, 41, 50, 27, 32, 41, 48, 60, 43)
#assuming we looked at sample and feel a t-test is appropriate, then we use it

t.test(otter_length)
#the base function of a t-test is comparing to a null hypothesis that the mean is equal to 0, so ignore p-value. We're just getting a CI
```

## T-test (1-sample t-test)

?t.test will show you how the function works in the help tab. Works for any function. We will change mu to dif settings (plus other changes)

See a claim: mean otter length is 50 inches

We have a sample, we decided its from a normally distributed blah blah

```{r}
otter_test <- t.test(x = otter_length, mu = 50)
otter_test
```

p-value of 0.0444, means that if the pop mean really is 50, there is a 4.44% chance that we could have taken a sample with the mean at least as different as 42.222

## Two-sample t-test

Are the differences in my two sample means so different that I think they are drawn from populations with different means.

```{r}
desert_bighorns <- c(32, 44, 18, 26, 50, 33, 42, 20)
#we need to have a big enough sample to apply central limit theory, so we have to assume we think the data is normally distributed. Does a t-dist make sense. We assume so for this.
sierra_bighorns <- c(28, 31, 40, 42, 26, 29, 31)

t.test(x = desert_bighorns, y = sierra_bighorns)
#p-value is how likely we could have drawn these samples from pops with the same mean. Low value, suggests maybe the pops were different. Lots of considerations though.
```

## Part 2: Two-sample t-test

How different 

```{r}
compact_suv <- mpg %>% 
  filter(class %in% c("compact", "suv"))
```

### Exploratory data visualization

Exploratory histograms:

```{r}
ggplot(data = compact_suv, aes(x = cty)) +
  geom_histogram(bins = 15) + #bins specified to help visualize, there are rules though
  facet_wrap(~class) #groups the hist by compact and suv
```

### Q-Q plot

Quantile - quantile plot, you want a linear relationship, closer you are is closer to a normally distributed data set. 

```{r}
ggplot(data = compact_suv, aes(sample = cty)) +
  geom_qq() +
  facet_wrap(~class) #they look close to linear, so because of the hists and the qq we think they're close to normally distributed
```

### Descriptive statistics

```{r}
car_stats <- compact_suv %>% 
  group_by(class) %>% 
  summarise(mean_city = mean(cty), 
            sd_city = sd(cty),
            n = n())
```

You should do all of these before considering a t-test. You also need to meet other assumptions (random, representative, etc.) Because we have met them, we think we can run a two sample t-test responsibly.

### 2-sample t-test

Create vectors containing observations for each sample. (city gas mileage)
```{r}
compact_sample <- compact_suv %>% 
  filter(class == "compact") %>% 
  pull(cty) #extracts data from a column and stores them as a vector, select would store it as a data frame and we need a vector

suv_sample <- compact_suv %>% 
  filter(class == "suv") %>% 
  pull(cty)

t.test(x = compact_sample, y = suv_sample)
```

